---
title: "NEON TOS Coding Lab Solution"
author: "K Duffy"
date: "8/19/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Loading the libraries that we need:
```{r, message=FALSE, warning=FALSE}
library(neonUtilities)
library(geoNEON)
library(sp)
library(dplyr)
library(ggplot2)
```

Pulling in the code referenced in the textbook:
```{r, message=FALSE, warning=FALSE, include=FALSE}
source('/Users/kdw223/Research/katharynduffy.github.io/neon_token_source.R')
veglist <- loadByProduct(dpID="DP1.10098.001", site="WREF", package="basic", check.size=FALSE, token = NEON_TOKEN)
summary(veglist)

```
```{r}
vegmap <- getLocTOS(veglist$vst_mappingandtagging, 
                          "vst_mappingandtagging")
```

```{r}
veg <- merge(veglist$vst_apparentindividual, vegmap, 
             by=c("individualID","namedLocation",
                  "domainID","siteID","plotID"))
```


Here's our original plot from the textbook:
```{r}
symbols(veg$adjEasting[which(veg$plotID=="WREF_075")], 
        veg$adjNorthing[which(veg$plotID=="WREF_075")], 
        circles=veg$stemDiameter[which(veg$plotID=="WREF_075")]/100/2, 
        inches=F, xlab="Easting", ylab="Northing")
```

**SOLUTIONS: NEON Part 1**

1. Convert the above diameter plot into a ggplot:
2. Set the color of your circles to be a function of each species:
```{r, message=FALSE, warning=FALSE}
ggplot(veg[which(veg$plotID=="WREF_075"),], aes(x=adjEasting, y=adjNorthing, color=taxonID, size=stemDiameter)) +
  geom_point()

```

3. Generate a histogram of tree heights for each plot.  Color your stacked bar as a function of each species:

First, I'll just make a general histogram:
```{r, message=FALSE}
#I'll remove NA values because they will mess up my model later in this code and give me a bunch of warnings:

veg=veg%>%
  filter(!is.na(stemDiameter))%>%
  filter(!is.na(height))

ggplot(veg, aes(x=height)) + 
  geom_histogram()

```
Now I'll add color as a function of species:

```{r, message=FALSE, warning=FALSE}
ggplot(veg, aes(x=height, fill=taxonID)) + 
  geom_histogram()

```

And now I will facet_wrap as a function of plotID:

```{r, message=FALSE, warning=FALSE}
ggplot(veg, aes(x=height, fill=taxonID)) + 
  geom_histogram()+
  facet_wrap(~ plotID)

```

4. Use `dplyr` to remove dead trees:

First I've looked at the data and metadata.  I'm going to figure out all the `unique` values of `plantStatus` at this site:

```{r}
unique(veg$plantStatus)
```
Given that check, it looks like I need to filter for "Dead, broken bole" and "Standing dead":

```{r}

vegLIVE=veg%>%
filter(plantStatus != c("Standing dead","Dead, broken bole"))

```

Now a couple of important and quick 'common sense' checks to make sure I really did filter out the dead trees:

First I'll just check how many rows I removed using the base function `nrow` (number of rows)
```{r}
nrow(veg)-nrow(vegLIVE)
```

Okay!  I removed >100 observations, but I'm getting an error message. Let's just check what all the `plantStatus` values look like now:
```{r}
unique(vegLIVE$plantStatus)
```

Oh no!  I still have dead trees!  This is *exactly* why we always check. 

From here we have a couple of options for filtering. I'll use a `string compare` function from `tidyverse`.  Also, I find `dplyr::filter` to be problematic when searching for multiple things at the same time, so I've broken those filters up for readability/ease:

```{r, message=FALSE, warning=FALSE,}
library(tidyverse)
vegLIVE=vegLIVE %>%
  filter(!str_detect(plantStatus, 'Standing dead'))%>%
  filter(!str_detect(plantStatus, 'Dead, broken bole'))
```

Let's do another common sense check:
```{r}
unique(vegLIVE$plantStatus)
```

No dead trees, success!

**To make things a little easier from here on out I'll just plot the top 3 species**

```{r}

vegTop3=vegLIVE%>%
  count(vegLIVE$taxonID) %>% 
  top_n(3)
summary(vegTop3)
```
Now I'll filter my main data.frame based on this top 3 data.frame:

```{r}
vegLIVE=vegLIVE%>%
  filter(is.element(vegLIVE$taxonID, vegTop3$`vegLIVE$taxonID`))
```

5. Create a simple linear model that uses DBH and height to predict allometries.  Print the summary information of your model:
```{r}
mdl=lm(height ~ stemDiameter * taxonID, data=vegLIVE) 
print(mdl)
```

Now I'll tack on my model predictions to my data.frame:
```{r, message=FALSE, warning=FALSE, include=FALSE}

vegLIVE <- vegLIVE %>%
  cbind( predict(mdl, newdata=., interval='confidence') )
summary(vegLIVE)
```
Let's take a look.  First let's plot our NEON data:

```{r, message=FALSE, warning=FALSE}
ggplot(data=vegLIVE, aes(x=stemDiameter, y=height, color=taxonID)) +
  geom_point()+
  facet_wrap(~ taxonID)
```
Now let's plot our model:
```{r, message=FALSE, warning=FALSE}

ggplot(data=vegLIVE, aes(x=stemDiameter, y=height, color=taxonID)) +
  geom_point() +
  geom_line( aes(y=fit) ) +
  geom_ribbon( aes( ymin=lwr, ymax=upr, fill=taxonID), alpha=.1 )+
  facet_wrap(~ taxonID)# alpha is the ribbon transparency
```

What do you think about our simple linear model?  To me it looks like we're failing to constrain the fact that trees have a height limit.  I guess gravity isn't just a 'theory'.  In this case I would certainy look at non-linear methods such as splines or a GAM, but we'll apply those methods later in this course.

7. Answer the following questions:

* How many `unique` species are present at `WREF`?

```{r}
print(length(unique(veg$taxonID)))
```
* What are the `top_5` trees based on height? 
```{r}

vegTop5_height=veg%>%
  slice_max(height, n=5)
head(vegTop5_height)
```

```{r}
vegTop5_height%>%
  select(plotID, plantStatus, taxonID, scientificName, adjNorthing, adjEasting, height)

```
Cool.  My native state tree, the Douglas Fir, dominates the height record (Go Oregon!).  I wonder if these trees are all together in an old growth stand, or if they're distributed throughout the area?

```{r}
ggplot(vegTop5_height, aes(x=adjEasting, y=adjNorthing, color=taxonID, size=height)) +
  geom_point(alpha=.4 )
```
It looks to me like we have a small old growth patch of Douglas Fir.  That would be super fun to study.

* Diameter?
```{r}
vegTop5_diameter=veg%>%
  slice_max(stemDiameter, n=5)
head(vegTop5_diameter)
```
```{r}
ggplot(vegTop5_diameter, aes(x=adjEasting, y=adjNorthing, color=taxonID, size=stemDiameter)) +
  geom_point(alpha=.4 )
```
Interesting.  Also mostly Douglas Firs, but not the ones who are the tallest.  I bet that's why our linear model failed.


* What proportion of sampled trees are dead?

```{r}
length(veg)/length(vegLIVE)
```

Or ~4%